---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bluebike

<!-- badges: start -->
<!-- badges: end -->

Bluebike contains data from [Blue Bikes System Data](https://www.bluebikes.com/system-data) website. It contains monthly trip data from 2020 to 2022. The data includes the name and geographic location of the bike stations, start and end time of the trip, trip duration, bike id, user information. Additionally, the data includes a random sample of 1000 trip history entries for sample analysis and functions for importation of the full data set.

## Installation

You can install the development version of bluebike from [GitHub](https://github.com/zyang2k/bluebike) with:

``` r
# install.packages("devtools")
devtools::install_github("zyang2k/bluebike")
```
## Sample Data Set

- `trip_history_sample`: a sample of 1000 trip data entries from February 2022.

## Data Wrangling

Count the number of trips from starting stations:

```{r example, message=FALSE, warning=FALSE}
library(bluebike)
library(dplyr)
stations <- trip_history_sample %>% 
  group_by(`start station name`) %>% 
  summarize(trips_from = n())
head(stations)
```

## Data Visualization

```{r, message=FALSE, warning=FALSE, eval= FALSE}
library(sf)
library(ggplot2)

start_station <- st_as_sf(trip_history_sample, coords = c('start station longitude', 'start station latitude'))

neighborhoods <- st_read("data-raw/Boston_Neighborhoods/Boston_Neighborhoods.shp")
neighborhoods_1 <- neighborhoods %>% st_set_crs(st_crs(polygon))


my_var <- highways$osm_lines %>% st_set_crs(st_crs(polygon))
                                            
ggplot() +
  geom_sf(data = my_var, aes(color=highway),
          size = .4,
          alpha = .65)+
  theme_void() +
  geom_sf(data = start_station, 
          size = .1,
          alpha = .1)

ggplot() +
  geom_sf(data = highways$osm_lines) 

ggplot(data = start_station) +
  geom_sf()
```


## Contributors
- [Ziyue Yang](https://github.com/zyang2k)
- [Tianshu Zhang](https://github.com/tianshu-zhang)
